---
title: "R basics"
author: "TA: Yuanyuan Li (yynli@ucdavis.edu)"
output: 
  pdf_document:
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=6, fig.height=4) 
```

# 1. Calculation: Use R as a calculator

```{r}
31/4*(37-25) 
3^2
sqrt(36)
log(4)
cos(6)
```

# 2. Objects 
## Assign values to object 'x' using any one of the following:
```{r}
x = 5 
x <- 5
5 -> x
```


## Calculation
```{r}
x+3
x^2
sqrt(x)
y = x^2
```

## Overwrite the existing objects
```{r}
x = 10
x = x+1
```


#3. Vectors
```{r}
# Create a vector
c(1,3,2,4)
# Save the vector as 'x'
x = c(1,3,2,4)

# R applies functions to every element of a vector
x - 10
x^2
```


#4. Some useful functions
```{r}
mean(x) # mean
sd(x) # standard deviation
var(x) # variance
summary(x)
sum(x) # sum of all elements
prod(x) # product of all elements
length(x) # number of elements

x[1:3] # the first three elements
```

#5. Import the data
* Use "Import Dataset" button in the menu of Environment: 
    + Read .txt file or .csv file : Import Dataset>From Text(base)>choose file>open>import 
    + Read Excel file : Import Dataset>From Excel>choose file>open>import 

* Read in a data set by specifying the full file path (remember to replace “\\” with “/” if your file path include “\\”): 
```{r}
data1 = read.table("~/Desktop/datasets/airfreight+breakage.txt")
```

* Read in a data set by setting the parent folder as working directory, then use "read.csv" or "read.table" function:
```{r}
setwd("~/Desktop/datasets")#set working directory to "datasets" folder
data1 = read.table("airfreight+breakage.txt")#read the data set in the folder
```

#6. Accessing specific rows and columns in a dataframe
```{r}
head(data1)#Display the first six rows
data1[1:3,]#Display the first six rows by row index
Y = data1[,1] # Extract variables from dataset
X = data1[,2]
data1[3,2]#Display the value is row 3 and column 2
```


#7. Plot 
```{r}
# histograms
hist(Y) 

# boxplots
boxplot(Y, main = 'Boxplot of Y')

# scatterplots
plot(X, Y, xlab = 'X', ylab = 'Y', main = 'Plot of Y versus X')
```



